\chapter{Introduction}\label{part:introduction}

This report covers the creation of a programming language for computer simulations. The language is constructed with scientists, who wishes to compute large simulations in different fields of science, as the primary user-group. Large simulations refers to problems which are deemed infeasible to be run on regular computers such as laptops and desktops. Focusing on such large simulations effectively gives the language a subfocus on parallel computing.

The aim of the language is not to compete with tools in the field of parallel computing, such as C and Fortran, which admittedly are very powerful and efficient programming languages, but are beginning to show their age and both lack a lot of prominent functionalities, such as pattern matching, for-each loops and arbitrary precision integers. 

Instead this language tries to provide the computing power of a supercomputer to scientists, who do not have extended experience with computer science. Currently this is a niche in the use of supercomputers and cluster computers, but considering the continued stride towards computers with a growing number of processor cores/units, we expect the accessibility of large parallel computers to increase. 

The Language Described in this Report (TLDR), aims to modernise this niche field and attempt to take over the reigns from the giants of yesteryear.

This premise gives the following problem statement:
\\

\emph{How can one define a programming language which allows a programmer to describe arbitrary scientific simulations, in a way that closely resembles formal mathematics and inherently allows for concurrent processing.}
