As described in \cref{\jenote{find afsnit om hvorfor vi skal køre parallelt}} most simulations are very heavy and need HPC to operate at resonable speeds. Furthermore HPC runs parallel to utilize the performance of multiple processors. On modern architectures parallelism is achieved via system threads. Threads can also be made concurrent if wanted but to achieve parallelism threads are needed. As described in \cref{sec:models}, models are used as the conceptional still images of simulations and general dynamic behaviour.  We therefore need a way to create models that can operate parallel. Threads are not easy to use to abstract over the definition of models. Since this language focuses heavily on models and simulations a more compatible abstraction is needed.

If we look at the definition given earlier about models they consist of a general theory and a description of objects and systems. This is a definition is very similar to how object oriented programming works, where all programs consists of objects, with fields and methods, and systems are build around these. It is though not surprising that this is a property of object oriented programming since it has been designed to model the real world.
Because of this it would be very beneficial to use a concurrent abstraction that has the same base where everything is objects.
The actor model is often described as the object oriented concurrency model. The next section will describe why this is the case, but because of its similarity to object oriented thinking the actor model was found to be the best concurrency model for simulations and therefore also The Language Described in this Report.